---
title: "Take-home Excersied 1"
site: distill::distill_website
---
# 1.0 Introduction

This is a visual analysis project which uses the data from [VAST Challenge 2022](https://vast-challenge.github.io/2022/). 
- Data: Participants.csv
Contains information about the residents of Engagement,OH that have a greed to participate in this study.
- Packages: tidyverse, ggdist, gghalves
The data is processed by using appropriate tidyverse family of packages and the statistical graphics is prepared using ggplot2 and its extensions.

```{r include=FALSE}
packages = c('tidyverse','ggdist','gghalves')
for(p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p, character.only = T)
}
```

# Read-in data
Firstly, we need to import our dataset. The source file is in csv format, and consists of . Hence, read_csv() of the readr package is used to import the data.It has 7 attributes.
- participantId: int
- householdSize: int
- haveKids(boolean):whether there arechildrenlivingintheparticipant’shousehold
- age: (integer):participant’s age in years at the start of the study 
- educationLevel(stringfactor):the participant’s education level,one of: {"Low","HighSchoolOrCollege","Bachelors","Graduate"}
- interestGroup(char):a char representing the participant’s stated primary interest
- joviality(float):avaluerangingfrom[0,1]indicatingtheparticipant’soverall
```{r layout="l-body-outset", preview=TRUE}
# Reading the csv file as a tbl_df
participants_data <- read_csv("./data/Participants.csv")
# Inspecting the structure of the dataset
# str(participants_data)
glimpse(participants_data)
```
```{r}
exam_data <- read_csv("data/Exam_data.csv")
ggplot(data=exam_data,
aes(x=RACE)) +
geom_bar()
ggplot(data=exam_data,
aes(x = MATHS)) +
geom_dotplot(dotsize = 0.5)
ggplot(data=exam_data,
aes(x = MATHS)) +
geom_dotplot(binwidth=2.5,
dotsize = 0.5) +
scale_y_continuous(NULL,
breaks = NULL)
ggplot(data=exam_data,
aes(x = MATHS)) +
geom_histogram()
ggplot(data=exam_data,
aes(x= MATHS,
fill = GENDER)) +
geom_histogram(bins=20,
color="grey30")
ggplot(data=exam_data,
aes(y = MATHS,
x= GENDER)) +
geom_boxplot()
ggplot(data=exam_data,
aes(y = MATHS,
x= GENDER)) +
geom_boxplot() +
geom_point(position="jitter",
size = 0.5)
ggplot(data=exam_data,
aes(y = MATHS,
x= GENDER)) +
geom_violin()
ggplot(data=exam_data,
aes(y = MATHS,
x= GENDER)) +
geom_violin(fill="light blue") +
geom_boxplot(alpha=0.5)
ggplot(data=exam_data,
aes(x= MATHS,
y=ENGLISH)) +
geom_point()
```

```{r layout="l-body-outset", preview=TRUE}
ggplot(data=participants_data, aes(x = age,fill=interestGroup)) +
geom_histogram(bins=20,
boundary = 100,
color="black",
fill="light blue") +
ggtitle("Distribution of Age")
```

```{r layout="l-body-outset", preview=TRUE}
ggplot(data=participants_data,
aes(y = age,
x= educationLevel)) +
geom_boxplot()
```

```{r layout="l-body-outset", preview=TRUE}
ggplot(data=participants_data,
aes(y = age,
x= interestGroup)) +geom_violin()
```