---
title: "In Class Exercise 7"
author:
- name: "Huang Anni"
  affiliation: "Singapore Management University"
description: |
  Geolocation data
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
---

```{r}
packages = c('igraph', 'tidygraph', 
             'ggraph', 'visNetwork', 
             'lubridate', 'clock',
             'tidyverse', 'ggmap',
             'rmarkdown')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

```{r}
GAStech_nodes <- read_csv("data/GAStech_email_node.csv")
GAStech_edges <- read_csv("data/GAStech_email_edge-v2.csv")
```
```{r}
glimpse(GAStech_edges)
```

The code chunk below will be used to perform the changes.
Things to learn from the code chunk above:

- both dmy() and wday() are functions of **lubridate** package. lubridate is an R package that makes it easier to work with dates and times.
- dmy() transforms the SentDate to Date data type.
- wday() returns the day of the week as a decimal number or an ordered factor if label is TRUE. The argument abbr is FALSE keep the daya spells in full, i.e. Monday. The function will create a new column in the data.frame i.e. Weekday and the output of wday() will save in this newly created field.
- the values in the Weekday field are in ordinal scale.

## Wrangling time
The code chunk below will be used to perform the changes.



```{r}
GAStech_edges$SentDate  = lubridate::dmy(GAStech_edges$SentDate)
GAStech_edges$Weekday = lubridate::wday(GAStech_edges$SentDate,label=TRUE, abbr = FALSE)
```

## Wrangling attributes
A close examination of GAStech_edges data.frame reveals that it consists of individual e-mail flow records. This is not very useful for visualisation.

In view of this, we will aggregate the individual by date, senders, receivers, main subject and day of the week.

Things to learn from the code chunk above:

four functions from dplyr package are used. They are: filter(), group(), summarise(), and ungroup().
The output data.frame is called GAStech_edges_aggregated.
A new field called Weight has been added in GAStech_edges_aggregated.

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  filter(MainSubject == "Work related") %>%
  group_by(source, target, Weekday) %>%
    summarise(Weight = n()) %>%
  filter(source!=target) %>%
  filter(Weight > 1) %>%
  ungroup()
```

## Reviewing the revised edges file

Table below shows the data structure of the reformatted GAStech_edges data frame

```{r}
paged_table(GAStech_edges_aggregated)
```
```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes,
                           edges = GAStech_edges_aggregated, 
                           directed = TRUE)
GAStech_graph
```
- You can **change which tibble data frame is active** with the activate() function. 
- If we wanted to rearrange the rows in the edges tibble to list those with the highest "weight" first, we could use activate() and then arrange().

```{r}
GAStech_graph %>%
  activate(edges) %>%
  arrange(desc(Weight))
```
```{r}
set.seed(1234)
ggraph(GAStech_graph,
       layout = "stress") + # random
  geom_edge_link() +
  geom_node_point() +
  theme_graph() # remove gray background

```
```{r}
set.seed(1234)
ggraph(GAStech_graph,
       layout = "stress") + # random
  geom_edge_link(aes(width=Weight,alpha=0.2)) +
  scale_edge_width(range = c(0.1,5))+
  geom_node_point(aes(color=Department, 
                      size = 3)) +
  theme_graph() # remove gray background

```

## Working with facet

```{r}
set.seed(1234)
ggraph(GAStech_graph,
       layout = "stress") + # random
  geom_edge_link(aes(width=Weight,alpha=0.2)) +
  scale_edge_width(range = c(0.1,5))+
  geom_node_point(aes(color=Department, 
                      size = 1)) +
  theme_graph() + # remove gray background
  facet_edges(~Weekday)
```
## Change the facet

```{r}
set.seed(1234)
ggraph(GAStech_graph,
       layout = "stress") + # random
  geom_edge_link(aes(width=Weight,alpha=0.2)) +
  scale_edge_width(range = c(0.1,5))+
  geom_node_point(aes(color=Department, 
                      size = 1)) +
  theme_graph() + # remove gray background
  facet_nodes(~Department)
```

## Network with centrality metric

```{r}
GAStech_graph %>%
  mutate(betweenness_centrality = centrality_betweenness()) %>%
  ggraph(
       layout = "stress") + # random
  geom_edge_link(aes(width=Weight,
                     alpha=0.2)) +
  scale_edge_width(range = c(0.1,5))+
  geom_node_point(aes(color=Department, 
                      size = betweenness_centrality)) +
  theme_graph() 
```
## Visualising Community

```{r}
g <- GAStech_graph %>%
  mutate(community = as.factor(group_edge_betweenness(weights = Weight, directed = TRUE))) %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = community))  
g + theme_graph()
```

## Building Interactive Network Graph with visNetwork

visNetwork() is a R package for network visualization, using vis.js javascript library.

visNetwork() function uses a nodes list and edges list to create an interactive graph.

The nodes list must include an "id" column, and the edge list must have "from" and "to" columns.
The function also plots the labels for the nodes, using the names of the actors from the "label" column in the node list.
The resulting graph is fun to play around with.

You can move the nodes and the graph will use an algorithm to keep the nodes properly spaced.
You can also zoom in and out on the plot and move it around to re-center it.

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  left_join(GAStech_nodes, by = c("sourceLabel" = "label")) %>%
  rename(from = id) %>%
  left_join(GAStech_nodes, by = c("targetLabel" = "label")) %>%
  rename(to = id) %>%
  filter(MainSubject == "Work related") %>%
  group_by(from, to) %>%
    summarise(weight = n()) %>%
  filter(from!=to) %>%
  filter(weight > 1) %>%
  ungroup()
visNetwork(GAStech_nodes, 
           GAStech_edges_aggregated)
```

## Working with layout

In the code chunk below, Fruchterman and Reingold layout is used.

```{r}
 visNetwork(GAStech_nodes,
            GAStech_edges_aggregated) %>%
   visIgraphLayout(layout = "layout_with_fr")
```

Visit Igraph to find out more about visIgraphLayout's argument.

## Working with visual attributes - Nodes

visNetwork() looks for a field called "group" in the nodes object and colour the nodes according to the values of the group field.

The code chunk below rename Department field to group.
```{r}
GAStech_nodes <- GAStech_nodes %>%
  rename(group = Department)
```

## Working with visual attributes - Nodes

When we rerun the code chunk below, visNetwork shades the nodes by assigning unique colour to each category in the group field.

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

## Working with visual attributes - Edges
In the code run below visEdges() is used to symbolise the edges.

- The argument arrows is used to define where to place the arrow.
- The smooth argument is used to plot the edges using a smooth curve.

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visEdges(arrows = "to", 
           smooth = list(enabled = TRUE, 
                         type = "curvedCW")) %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```