---
title: "In Class Exercise 8"
author:
- name: "Huang Anni"
  affiliation: "Singapore Management University"
description: |
  Geolocation data
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
---

```{r}
packages = c('tidyverse', 'lubridate', 
             'ggthemes', 'reactable', 
             'reactablefmtr',
             'gt', 'gtExtras')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```
# Data Preparation

## Importing Data
The code chunk below is used to import CoffeeChain.rds into R.

```{r import-data}
coffeechain <- read_rds("data/CoffeeChain.rds")
```

## Data wrangling
The code chunk below is used to aggregate Sales and Budgeted Sales at the Product level.

```{r aggregate-at-product-level}
product <- coffeechain %>%
  group_by(`Product`) %>%
  summarise(`target` = sum(`Budget Sales`),
            `current` = sum(`Sales`)) %>%
  ungroup()
```

# Bullet chart in ggplot2

```{r bullet-chart}
ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01),
           fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75),
           fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5),
           fill="grey50", width=0.85) +
  geom_col(aes(Product, current), 
           width=0.35,
           fill = "black") + 
  geom_errorbar(aes(y = target,
                    x = Product, 
                    ymin = target,
                    ymax= target), 
                width = .4,
                colour = "red",
                size = 1) +
  coord_flip()
```

# Plotting sparklines using ggplot2


In this section, you will learn how to plot sparklines by using ggplot2.

## Preparing the data

```{r sparkline-data-prep}
sales_report <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  mutate(Month = month(Date)) %>%
  group_by(Month, Product) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup() %>%
  select(Month, Product, Sales)
```

The code chunk below is used to compute the minimum, maximum and end othe the month sales.

```{r compute-stats-sparkline}
mins <- group_by(sales_report, Product) %>% 
  slice(which.min(Sales))
maxs <- group_by(sales_report, Product) %>% 
  slice(which.max(Sales))
ends <- group_by(sales_report, Product) %>% 
  filter(Month == max(Month))
```

The code chunk below is used to compute the 25 and 75 quantiles.

```{r}
quarts <- sales_report %>%
  group_by(Product) %>%
  summarise(quart1 = quantile(Sales, 
                              0.25),
            quart2 = quantile(Sales, 
                              0.75)) %>%
  right_join(sales_report)
```

# sparklines in ggplot2

```{r ggplot-sparkline}
ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, aes(ymin = quart1, max = quart2), 
              fill = 'grey90') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), hjust = 0, nudge_x = 0.5) +
  geom_text(data = ends, aes(label = Product), hjust = 0, nudge_x = 1) +
  expand_limits(x = max(sales_report$Month) + 
                  (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  theme_tufte(base_size = 3, base_family = "Helvetica") +
  theme(axis.title=element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), strip.text = element_blank())
```

# Static Information Dashboard Design: gt and gtExtras methods

In this section, you will learn how to create static information dashboard by using [gt](https://gt.rstudio.com/index.html) and [gtExtras](https://jthomasmock.github.io/gtExtras/index.html) packages. Before getting started, it is highly recommended for you to visit the webpage of these two packages and review all the materials provided on the webpages at least once. You done not have to understand and remember everything provided but at least have an overview of the purposes and functions provided by them.

# Plotting a simple bullet chart

In this section, you will learn how to prepare a bullet chart report by using functions of gt and gtExtras packages.

```{r create-simple-bullet-chart}
product %>%
  gt::gt() %>%
  gt_plt_bullet(column = current, 
              target = target, 
              width = 60,
              palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```

## sparklines: gtExtras method

Before we can prepare the sales report by product by using gtExtras functions, code chunk below will be used to prepare the data.

```{r gtExtras-func}
report <- coffeechain %>%
  mutate(Year = year(Date)) %>%
  filter(Year == "2013") %>%
  mutate (Month = month(Date, 
                        label = TRUE, 
                        abbr = TRUE)) %>%
  group_by(Product, Month) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup()
```

It is important to note that one of the requirement of gtExtras functions is that almost exclusively they require you to pass data.frame with list columns. In view of this, code chunk below will be used to convert the report data.frame into list columns.

```{r change-dataframe-into-list-columns}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```

## Plotting Coffechain Sales report

```{r coffechain-sales-report}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop") %>%
   gt() %>%
   gt_plt_sparkline('Monthly Sales')
```

## Adding statistics

First, calculate summary statistics by using the code chunk below.

```{r add-stats-to-coffe-sales-report}
report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            ) %>%
  gt() %>%
  fmt_number(columns = 4,
    decimals = 2)
```

## Combining the data.frame

Next, use the code chunk below to add the statistics on the table.

```{r change-into-monthly-sales-list}
spark <- report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
spark
```

```{r summarise-min-max-per-product}
sales <- report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            )
sales
```

```{r join-with-original-data}
sales_data = left_join(sales, spark)
sales_data
```

## Plotting the updated data.table

```{r updated-datatable}
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales')
```

## Combining bullet chart and sparklines

Similarly, we can combining the bullet chart and sparklines using the steps below.

```{r combine-bullet-chart-and-sparklines}
bullet <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  group_by(`Product`) %>%
  summarise(`Target` = sum(`Budget Sales`),
            `Actual` = sum(`Sales`)) %>%
  ungroup()
```

```{r left-join-bullet-data}
sales_data = sales_data %>%
  left_join(bullet)
```

```{r plot-bullet-chart-and-sparklines}
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales') %>%
  gt_plt_bullet(column = Actual, 
                target = Target, 
                width = 28,
                palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```

# Interactive Information Dashboard Design: reactable and reactablefmtr methods

In this section, you will learn how to create interactive information dashboard by using [reactable](https://glin.github.io/reactable/index.html) and [reactablefmtr](https://kcuilla.github.io/reactablefmtr/index.html) packages. Before getting started, it is highly recommended for you to visit the webpage of these two packages and review all the materials provided on the webpages at least once. You done not have to understand and remember everything provided but at least have an overview of the purposes and functions provided by them.

In order to build an interactive sparklines, we need to install dataui R package by using the code chunk below.

```{r install-dataui}
remotes::install_github("timelyportfolio/dataui")
```

Next, you all need to load the package onto R environment by using the code chunk below.

```{r import dataui}
library(dataui)
```

## Plotting interactive sparklines

Similar to gtExtras, to plot an interactive sparklines by using reactablefmtr package we need to prepare the list field by using the code chunk below.

```{r interactive-sparklines}
report <- report %>%
  group_by(Product) %>%
  summarize(`Monthly Sales` = list(Sales))
```

Next, react_sparkline will be to plot the sparklines as shown below.

```{r}
reactable(
  report,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```

## Changing the pagesize

By default the pagesize is 10. In the code chunk below, arguments defaultPageSize is used to change the default setting.

```{r change-pagesize}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```

## Adding pointa and labels

In the code chunk below highlight_points argument is used to show the minimum and maximum values points and label argument is used to label first and last values.

```{r add-point-labels}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        labels = c("first", "last")
        )
    )
  )
)
```

## Adding reference line

In the code chunk below statline argument is used to show the mean line.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        statline = "mean"
        )
    )
  )
)
```

## Adding bandline

Instead adding reference line, bandline can be added by using the bandline argument.

```{r add-bandline}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        line_width = 1,
        bandline = "innerquartiles",
        bandline_color = "green"
        )
    )
  )
)
```

## Changing from sparkline to sparkbar

Instead of displaying the values as sparklines, we can display them as sparkbars as shiwn below.

```{r change-from-sparkline-to-sparkbar}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkbar(
        report,
        highlight_bars = highlight_bars(
          min = "red", max = "blue"),
        bandline = "innerquartiles",
        statline = "mean")
    )
  )
)
```
