---
title: "Hands-on excercise 9"
description: |
  Analysing financial data
author:
  - name: Huang Anni 
    url: https://huanganni.netlify.app/huanganni.html
    affiliation: "Singapore Management University"
    affiliation_url: https://www.smu.edu.sg/
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
---

```{r install-packages}
packages = c('tidyverse', 'lubridate', 
             'tidyquant', 'ggHoriPlot',
             'timetk', 'ggthemes',
             'plotly')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```
# Data Preparation

## Importing Data
The code chunk below is used to import companySG.csv into R.

```{r import-data}
company <- read_csv("./In-class_Ex10/data/companySG.csv")
```

## Data wrangling

tidyquant integrates resources for collecting and analysing financial data with the tidy data infrastructure of the tidyverse, allowing for seamless interaction between each.

In this section, you will learn how to extract the daily stock values of a selected stocks from Yahoo Finance by using tidyquant.

### Step 1: Pre-processing

We will import a pre-prepared company list called companySG.csv onto R. The list consists of top 45 companies by market capitalisation in Singapore. However, we just want the top 40.

```{r top-40}
Top40 <- company %>% 
  slice_max(`marketcap`, n=40) %>% 
  select(symbol)
```

### Step 2: Get specific stock

tq_get() method will be used to extract daily values of these stocks from [Yahoo Finance](https://sg.finance.yahoo.com/?guccounter=1&guce_referrer=aHR0cHM6Ly93d3cuZ29vZ2xlLmNvbS8&guce_referrer_sig=AQAAAHa98gxRGd7UR5U5CBJQrzOlGe7GebuoJHwAnhfZteZFhvBCGYuLXElRR4e1jYMKNVAfoTkxZywJqXNn70g1lsXCL4GV5S1iHwV4MK3pKg8H33FHqCteh3jXuTKvpx1sZCOe_tlooJkLH9Odhgj1OV72vk3eRsCgzUeGb6zDxqOe) via APIs. The time period for the data was set from 1st January 2020 to 31st March 2021. The data are specified to be returned in daily intervals.

```{r stock40daily}
Stock40_daily <- Top40 %>%
  tq_get(get = "stock.prices", 
         from = "2020-01-01", 
         to = "2022-03-31") %>%
  group_by(symbol) %>%
  tq_transmute(select = NULL, 
               mutate_fun = to.period, 
               period  = "days")
```

## Plotting a horizon graph

In this section, you will learn how to plot a [horizon graph](http://www.perceptualedge.com/articles/visual_business_intelligence/time_on_the_horizon.pdf) by using [geom_horizon()](https://rivasiker.github.io/ggHoriPlot/reference/geom_horizon.html) of [ggHoriPlot](https://rivasiker.github.io/ggHoriPlot/index.html) package.

```{r horizon-graph}
Stock40_daily %>% 
  ggplot() +
  geom_horizon(aes(x = date, y=adjusted), origin = "midpoint", horizonscale = 6)+
  facet_grid(symbol~.)+
  theme_few() +
  scale_fill_hcl(palette = 'RdBu') +
  theme(panel.spacing.y=unit(0, "lines"), strip.text.y = element_text(
    size = 5, angle = 0, hjust = 0),
    legend.position = 'none',
    axis.text.y = element_blank(),
    axis.text.x = element_text(size=7),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    panel.border = element_blank()
    ) +
  scale_x_date(expand=c(0,0), date_breaks = "1 month", date_labels = "%b%y") +
  ggtitle('Daily Adjusted Prices (Jan 2020 to Mar 2022)')
```

## Horizon graph makeover

- Instead of showing stock code, the stock name will be displayed.
- Adding reference lines

### Step 1: Left join company data

left_join() of dplyr package is used to append fields from company data.frame onto Stock_daily data.frame. Next select() is used to select columns 1 to 8 and 11 to 12.

```{r}
Stock40_daily <- Stock40_daily %>%
  left_join(company) %>%
  select(1:8, 11:12)
```

### Step 2: Add the vertical reference lines

geom_vline() is used to add the vertical reference lines.

```{r}
Stock40_daily %>% 
  ggplot() +
  geom_horizon(aes(x = date, y=adjusted), origin = "midpoint", horizonscale = 6)+
  facet_grid(Name~.)+
  geom_vline(xintercept = as.Date("2020-03-11"), colour = "grey15", linetype = "dashed", size = 0.5)+
  geom_vline(xintercept = as.Date("2020-12-14"), colour = "grey15", linetype = "dashed", size = 0.5)+
  theme_few() +
  scale_fill_hcl(palette = 'RdBu') +
  theme(panel.spacing.y=unit(0, "lines"),
        strip.text.y = element_text(size = 5, angle = 0, hjust = 0),
    legend.position = 'none',
    axis.text.y = element_blank(),
    axis.text.x = element_text(size=7),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    panel.border = element_blank()
    ) +
  scale_x_date(expand=c(0,0), date_breaks = "1 month", date_labels = "%b%y") +
  ggtitle('Daily Adjusted Prices (Jan 2020 to Mar 2022)')
```
## Plotting Stock Price Line Graph: ggplot methods

In the code chunk below, geom_line() of ggplot2 is used to plot the stock prices.

```{r}
Stock40_daily %>%
  filter(symbol == "DBSDF") %>%
  ggplot(aes(x = date, y = close)) +
    geom_line() +
    labs(title = "DBS Group Holdings Ltd (DBSDF) Daily Stock Price", 
         y = "Closing Price", x = "") + 
    theme_tq()
```
## Plotting interactive stock price line graphs

In this section, we will create interactive line graphs for four selected stocks.

Step 1: Selecting the four stocks of interest.

```{r}
selected_stocks <-  Stock40_daily %>%
  filter (`symbol` == c("C09.SI", "SINGF", "SNGNF", "C52.SI"))
```

Step 2: Plotting the line graphs by using ggplot2 functions and ggplotly() of plotly R package

```{r}
p <- ggplot(selected_stocks, aes(x = date, y = adjusted))+
        scale_y_continuous() +
        geom_line() +
        facet_wrap(~Name, scales = "free_y",) +
        theme_tq() +
        labs(title = "Daily stock prices of selected weak stocks", x = "", y = "Adjusted Price") +
        theme(axis.text.x = element_text(size = 6), axis.text.y = element_text(size = 6))
ggplotly(p)
```
## Plotting Candlestick Chart: tidyquant method

In this section, you will learn how to plot [candlestick](https://en.wikipedia.org/wiki/Candlestick_chart) chart by using [geom_candlestick()](https://business-science.github.io/tidyquant/reference/geom_chart.html) of tidyquant package.

Before plotting the candlesticks, the code chunk below will be used to define the end data parameter. It will be used when setting date limits throughout the examples.

```{r}
end <- as_date("2022-03-31")
```

Now we are ready to plot the candlesticks by using the code chunk below.

```{r}
Stock40_daily %>%
  filter(symbol == "DBSDF") %>%
  ggplot(aes(
    x = date, y = close)) +
  geom_candlestick(aes(
    open = open, high = high, 
    low = low, close = close)) +
  geom_line(size = 0.5)+
    coord_x_date(xlim = c(end - weeks(12), 
                          end),
                 ylim = c(20, 35),
                 expand = TRUE) +
  labs(title = "DBS Group Holdings Ltd (DBSDF) Daily Stock Price", 
       y = "Closing Price", x = "") + 
  theme_tq()
```
## Plotting candlestick chart and MA lines: tidyquant method

Things to learn from the code chunk:

- [geom_MA](https://business-science.github.io/tidyquant/reference/geom_ma.html) is used to add the moving average line. It is a wrapper function of [SMA()](https://www.rdocumentation.org/packages/TTR/versions/0.24.3/topics/SMA) from the TTR package.

```{r}
Stock40_daily %>%
  filter(symbol == "DBSDF") %>%
  ggplot(aes(
    x = date, y = close)) +
  geom_candlestick(aes(
    open = open, high = high, 
    low = low, close = close)) +
  geom_line(size = 0.5)+
  geom_ma(color = "darkgreen", n = 20) +
  geom_ma(color = "lightgreen", n = 20) + 
    coord_x_date(xlim = c(end - weeks(12), 
                          end),
                 ylim = c(20, 35),
                 expand = TRUE) +
  labs(title = "DBS Group Holdings Ltd (DBSDF) Daily Stock Price",
       subtitle = "darkgreen = 1-day MA, lightgreen = 5-day MA",
       y = "Closing Price", x = "") + 
  theme_tq()
```

## Plotting Bollinger Bands: tidyquant method

In this section, you will learn how to plot [bollinger bands](https://en.wikipedia.org/wiki/Bollinger_Bands) by using [geom_bbands()](https://business-science.github.io/tidyquant/reference/geom_bbands.html) of tidyquant package.

Things you can learn from the code chunk:

- geom_bbands() plots a range around a moving average typically two standard deviations up and down. The moving average functions used are specified in SMA() from the TTR package.

```{r}
Stock40_daily %>%
  filter(symbol == "DBSDF") %>% 
  ggplot(aes(x=date, y=close))+
  geom_line(size=0.5)+
  geom_bbands(aes(
    high = high, low = low, close = close), 
    ma_fun = SMA, sd = 2, n = 20,
    size = 0.75, color_ma = "royalblue4", 
    color_bands = "red1")+
    coord_x_date(xlim = c("2020-02-01", 
                          "2022-03-31"), 
                 expand = TRUE)+
    labs(title = "DBS Group Holdings Ltd (DBSDF) Daily Stock Price",
         subtitle = "dotted red lines = bollinger bands",
         x = "Date", y ="Price") +
theme(legend.position="none")
```

## Plotting Interactive Candlesticks Chart: ggplot2 and plotly R method

First, a candleStick_plot function is written as follows:

Credit: I learned this trick from [RObservations #12: Making a Candlestick plot with the ggplot2 and tidyquant packages](https://www.r-bloggers.com/2021/09/robservations-12-making-a-candlestick-plot-with-the-ggplot2-and-tidyquant-packages/)

```{r}
candleStick_plot<-function(symbol, from, to){
  tq_get(symbol, from = from, to = to, warnings = FALSE) %>% 
    mutate(greenRed=ifelse(open-close>0, "Red", "Green")) %>% 
    ggplot()+
    geom_segment(aes(x = date, xend=date, y =open, yend =close, colour=greenRed), size=3)+
    theme_tq()+
    geom_segment(aes(x = date, xend=date, y =high, yend =low, colour=greenRed))+
    scale_color_manual(values=c("ForestGreen","Red"))+
    ggtitle(paste0(symbol," (",from," - ",to,")"))+
    theme(legend.position ="none",
          axis.title.y = element_blank(),
          axis.title.x=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),
          plot.title= element_text(hjust=0.5))
}
```

## Plotting interactive candlestick chart: ggplot2 and plotly R method

```{r}
p <- candleStick_plot("DBSDF",
                      from = '2022-01-01',
                      to = today())
ggplotly(p)
```

