---
title: "Hand-out 2"
site: distill::distill_website
---
```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
                      eval=TRUE,
                      warning=FALSE,
                      message=FALSE)
```

# Import packaged
```{r}
packages = c('tidyverse', 'ggdist', 'ggridges',
             'patchwork', 'ggthemes', 'hrbrthemes',
             'ggrepel', 'ggforce')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

# Read-in data

- The code chunk below imports exam_data.csv into R environment using read_csv() function of readr package.

- readr is one of the tidyverse package.

- Year end examination grades of a cohort of primary 3 students from a local school.

- There are a total of seven attributes. Four of them are categorical data type and the other three are in continuous data type.
  - The categorical attributes are: ID, CLASS, GENDER and RACE.
  - The continuous attributes are: MATHS, ENGLISH and SCIENCE.
```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Beyond ggplot2 Annotation
One of the challenge in plotting statistical graph is annotation, especially with large number of data points.
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

**ggrepel** is an extension of ggplot2 package which provides geoms for ggplot2 to repel overlapping text as in our examples on the right. We simply replace geom_text() by geom_text_repel() and geom_label() by geom_label_repel.
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

```
# Beyond ggplot2 Themes
ggplot2 comes with eight built-in themes, they are: theme_gray(), theme_bw(), theme_classic(), theme_dark(), theme_light(), theme_linedraw(), theme_minimal(), and theme_void().

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores")
```

