---
title: "Take-home Excercise 1"
site: distill::distill_website
---
# 1.0 Introduction

This is a visual analysis project which uses the data from [VAST Challenge 2022](https://vast-challenge.github.io/2022/). 

- **Data: Participants.csv**: Contains information about the residents of Engagement,OH that have a greed to participate in this study.

- Packages: tidyverse, ggdist, gghalves
The data is processed by using appropriate **tidyverse** family of packages and the statistical graphics is prepared using **ggplot2** and its **extensions**.

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
                      eval=TRUE,
                      warning=FALSE,
                      message=FALSE)
```

```{r layout="l-body-outset", preview=TRUE}
packages = c('tidyverse','ggdist','gghalves')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# Read-in data

Firstly, we need to import our dataset. The source file is in csv format, and consists of . Hence, read_csv() of the readr package is used to import the data.It has 7 attributes.

- participantId: int
- householdSize: int
- haveKids(boolean):whether there arechildrenlivingintheparticipant’shousehold
- age: (integer):participant’s age in years at the start of the study 
- educationLevel(stringfactor):the participant’s education level,one of: {"Low","HighSchoolOrCollege","Bachelors","Graduate"}
- interestGroup(char):a char representing the participant’s stated primary interest
- joviality(float):a value ranging from[0,1] indicating the participant’s overall happiness level at the start of the study
```{r layout="l-body-outset", preview=TRUE}
# Reading the csv file as a tbl_df
participants_data <- read_csv("./data/Participants.csv")
# Inspecting the structure of the dataset
# str(participants_data)
str(participants_data)
```

```{r}
# Convert speed_vector to ordered factor vector
education_factor <- factor(participants_data['educationLevel'],ordered=TRUE,levels=c("Low","HighSchoolOrCollege","Bachelors","Graduate"))
```

```{r layout="l-body-outset", preview=TRUE}
ggplot(data=participants_data,
aes(y = age,
x= educationLevel)) +
geom_boxplot()
```
Violin plot for 
```{r layout="l-body-outset", preview=TRUE}
ggplot(data=participants_data,
aes(y = age,x= interestGroup)) +geom_violin()+
ggtitle("Distribution of Age in different Interest Group")
```